### 2022-03-29 Writing/Palaeontology/Reviews/response_to_reviewers.tex

\setlength\parindent{0pt}

\begin{document}

\textbf{RE: PALA-11-19-4635-OA}\\
\bigskip
Dear Dr Benson,\\
\bigskip

We are very grateful to the three referees for their helpful and constructive comments, which we believe have helped me to significantly improve the manuscript.
Unfortunately, due to circumstances out of our control (pandemic, moving continent and new job), it took us way longer to reply to the reviewers requests as we could have done.
Fortunately, however, we have taken all of their comments on board, and respond to their points in details below.

@@

\textcolor{blue}{My example points:}

\begin{itemize}

\item{\textcolor{blue}{The abstract lists "shared phylogenetic history" as a factor that causes character correlation in a list of three factors. Approximately the same list of factors is presented in the introduction (p.5, bulletted), where "evolutionary dependence" is stated in place of "shared phylogenetic history". This seems confusing as the two lists seems to serve the same purpose but have non-overalpping content. Shared evolutionary history is a positive thing (we're trying to infer evolutionary history) whereas "evolutionary dependence" of traits seems like it could be negative in some circumstances (for the body elongation example, it may cause it to infer a clade of long-bodied taxa).}}

%@@@TODO-1: add a line in the intro explaining this? 
%TG: something like that "Note that investigating the shared evolutionary history is often part of the goal of phylogenetic analyses."

@@

\item{\textcolor{blue}{Matrices that were modified to minimise character differences are referred to as 'minimised'. However, in the context of the main questions, 'minimisation' of character differences corresponds to maximisation of character state correlations. Since these two concepts have opposite polarities, it would be easier for readers to follow the work if it consistently referred to one or the other - and I'd propose that referring to the amount of character correlation would be best and more straightforward for readers.}}

We have changed the naming of the matrices and the trees throughout the whole manuscript to now only focus on the character correlation aspect (\textit{c.f.} the character difference). The \textit{Normal} matrix/tree is now the \textit{Unperturbed} matrix/tree the \textit{Maximised (difference)} matrix/tree is now the \textit{Minimised (correlation)} matrix/tree and the \textit{Minimised (difference)} matrix/tree is now the \textit{Maximised (correlation)} matrix. The \textit{Randomised} matrix/tree stayed unchanged.

@@

\item{\textcolor{blue}{'Normal' has multiple meanings and mathematically could mean 'orthogonal'. In the way the authors use it, it comes across as colloquial, and it doesn't contain much information (there is no objective meaning of 'normal' in this context). A more intuitive term might be 'unmodified' or 'simnulated', or something else, there are probably better options.}}

We have now changed mentions to the \textit{Normal} tree as "the topology inferred from the unperturbed matrix" to avoid any confusion throughout the manuscript.

@@

\item{\textcolor{blue}{'In a few places 'phylogenetic signal' is used to refer to a situation where a topology is well-supported. But I had the impression this might not strictly correspond to support of the "true" phylogeny (correct me if I'm wrong however). Phylogenetic signal is a widely-used phrase with several different potential meanings. So to avoid confusion it would be useful each time if the text referred explictily to support for the "true" phylogeny underlying the simulations, or to support for nodes in the inferred phylogeny, whether they are present in the true phylogeny or not. Or tended to produce strong support for incorrect topologies.}}

We have now replaced mentions of "phylogenetic signal" to "\textit{phylogenetic information}" throughout the whole manuscript to avoid any confusion with other uses of the term "phylogenetic signal". Here we mean information (code) used to infer phylogenies.

@@

\section{Referee 2}

\begin{itemize}

\item{\textcolor{blue}{Lines 22-23: 'probabilistic methods can easily accommodate for a random correlation between characters.'
\\
'random correlation' sounds like an oxymoron! I think that what the authors really mean is moderate amounts of correlation that basically mimic accelerated local rates of change on a small proportion of branches. From what the authors state later, it’s not the randomness that is important, but that it occurs infrequently relative to total taxon sampling.}}

We have now changed "random correlation" to "\textit{randomised} correlation" throughout the manuscript to better reflect this comment.

### 2022-03-29 Writing/Palaeontology/draft.tex
Editors comments sorted (missing 1)

\newpage
\begin{abstract}
Phylogenetic analysis algorithms assume character independence - a condition generally acknowledged to be violated by morphological data.
Correlation between characters can originate from intra-organismal features, \textcolor{blue}{evolutionary dependence} or forced by particular character-state coding schemes.
Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be diminished \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
In this study, we use a metric of raw character difference as a proxy for character correlation.
Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
This means that datasets with low correlation between characters will make it more complicated to estimate a correct topology.
However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing or increasing character correlation is not more effective in getting the expected topology than randomly drawing characters.
Furthermore, in bigger matrices (350 characters), the inference method has a strong effect with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
However, encouragingly, they also suggest that, unless correlation is actively maximised or minimised, probabilistic methods can easily accommodate for a random correlation between characters.

@@

We standardised each character by arbitrarily modifying their character state tokens (or symbols) by \textcolor{blue}{order of appearance in the matrix - for example for the first character (1st column) in the matrix, if the state coded for the first OTU (1st row) was \texttt{3}, it was translated into \texttt{1} for the whole character (across the 1st column); if the state coded for the second OTU (2nd row) was \texttt{0}, it was translated into \texttt{2} for the whole character, etc.}
We replaced all the occurrences of the first token to be \texttt{1}, the second to be \texttt{2}, etc.
This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; as used in \citealt{Davalos01072014}).
This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \textcolor{blue}{(see supplementary materials 1)}.
When the character difference is null ($0$) it means that characters convey the same phylogenetic signal (i.e. characters are entirely correlated). 
When the character difference is maximal ($1$) it means the characters convey the greatest difference in phylogenetic signal

@@

\subsection{Effect of character differences on topology}
We find that there is a significant influence of morphological character correlation on inferring tree topology.
When characters are correlated, matrices convey a strong (but potentially misleading) phylogenetic signal since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of characters giving a conflicting signal (homoplasy).
Throughout our simulation, the maximised character correlation scenario (i.e. decreased character difference) tended to result in trees slightly closer to the tree inferred from the unperturbed matrix and the minimised character correlation led to slightly worse trees (Fig. \ref{Fig:RF_results_best}).
These results suggest that, in matrices with low character correlations, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.

>>>>

\newpage
\begin{abstract}
Phylogenetic analysis algorithms assume character independence - a condition generally acknowledged to be violated by morphological data.
Correlation between characters can originate from intra-organismal features, \textcolor{blue}{evolutionary dependence} or forced by particular character-state coding schemes.
Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be diminished \textit{a priori} with careful character coding, phylogenetic software do not distinguish between any of them.
In this study, we use a metric of raw character difference as a proxy for character correlation.
Using thorough simulations, we test the effect of increasing or decreasing character differences on tree topology.
Overall, we found a negative effect of reducing character correlations on recovering the correct topology.
This means that datasets with low correlation between characters will make it more complicated to estimate a correct topology.
However, this effect is less important for matrices with a small number of taxa (25 in our simulations) where reducing or increasing character correlation is not more effective in getting the expected topology than randomly drawing characters.
Furthermore, in bigger matrices (350 characters), the inference method has a strong effect with Bayesian trees being consistently less affected by character correlation than maximum parsimony trees.
These results suggest that ignoring the problem of character correlation or independence can often impact topology in phylogenetic analysis.
However, encouragingly, they also suggest that, unless correlation is actively maximised or minimised, probabilistic methods can easily accommodate for a \textcolor{blue}{randomised} correlation between characters.

@@

We standardised each character by arbitrarily modifying their character state tokens (or symbols) by \textcolor{blue}{order of appearance in the matrix - for example for the first character (1st column) in the matrix, if the state coded for the first OTU (1st row) was \texttt{3}, it was translated into \texttt{1} for the whole character (across the 1st column); if the state coded for the second OTU (2nd row) was \texttt{0}, it was translated into \texttt{2} for the whole character, etc.}
We replaced all the occurrences of the first token to be \texttt{1}, the second to be \texttt{2}, etc.
This procedure allows comparison of characters regardless of the significance given to their tokens (following the \textit{xyz} notation in \citealt{felsenstein2004inferring}; as used in \citealt{Davalos01072014}).
This way, a character \texttt{A = \{2,2,3,0,0,3\}} for six taxa would be standardised as \texttt{A' = \{1,1,2,3,3,2\}} \textcolor{blue}{(see supplementary materials 1)}.
When the character difference is null ($0$) it means that characters convey the same \textcolor{blue}{phylogenetic information} (i.e. characters are entirely correlated). 
When the character difference is maximal ($1$) it means the characters convey the greatest difference in \textcolor{blue}{phylogenetic information}.

@@

\subsection{Effect of character differences on topology}
We find that there is a significant influence of morphological character correlation on inferring tree topology.
When characters are correlated, matrices convey a strong (but potentially misleading) \textcolor{blue}{phylogenetic information} since characters are more likely to agree with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker \textcolor{blue}{phylogenetic information}with a high amount of characters giving a conflicting signal (homoplasy).
Throughout our simulation, the maximised character correlation scenario (i.e. decreased character difference) tended to result in trees slightly closer to the tree inferred from the unperturbed matrix and the minimised character correlation led to slightly worse trees (Fig. \ref{Fig:RF_results_best}).
These results suggest that, in matrices with low character correlations, bootstrapping or other data randomisation methods can effectively improve the ability to recover the correct topology.
